---
output:
  pdf_document: default
  html_document: default
  word_document: default
---

Q1.

```{r}
library(ggplot2)
library(ggthemes)
library(alr4)
library(dplyr)
```

```{r}
help(cloud)
head(cloud)
```


From the link "https://rdrr.io/rforge/alr4/man/cloud.html"
I found the following description;
Description
The data summarize the results of the first Florida Area Cumulus Experiment, or FACE-1, designed to study the effectiveness of cloud seeding to increase rainfall in a target area (Woodley, Simpson, Biondini, and Berkley, 1977).

Format
This data frame contains the following columns:

A
Action, 1=seed, 0=do not seed

D
Day after June 16, 1975

S
Suitability for seeding

C
percent cloud cover in experimental area, measured using radar in Coral Gables, Florida

P
10^7 m^3 prewetness

E
echo motion category, either 1 or 2, a measure for type of cloud

Rain
10^7 m^3 in target area


```{r}
cloud$seed <- 1*(cloud$A = 1)
cloud$Noseed <- 1*(cloud$A = 0)

cloud$Seedexist[cloud$seed==1] <- 'Yes'
cloud$Seedexist[cloud$Noseed==1] <- 'No'
cloud$seedexist <- factor(cloud$Seedexist, levels = c('Yes','No'))
cloud$seedexist
scatterplotMatrix(cloud[,c(2:7)])

```

Observe that the 4th column "C" of the dataset cloud can effect the 7-th column "Rain" as exponential graph, as we can see on the (6,3) component of the avoce matrix plot.
(For the "C" column, the 2nd row has the value as 37.9 and thus this row is the outlier.)
So we can transfrom the "Rain" in the graph of the fitting of linear regression with x-axis "C" and y-axis "Rain" after grouping acoording to the existance of seeding(1st colum "A").


```{r}
cloud$LogRain <- log (cloud$Rain) 
summary(model1 <- lm(LogRain ~S , data=cloud))

#effect plot:
summary(cloud$LogRain)

ggplot( data = cloud , aes(x=S, y=LogRain )) + geom_line()
```
As the linear regression fitting line, we can see the coefficient of the above linear model regression has nonzero intercept(constant) "b" in the linear regression of LogRain = a*S + b with the t-test result in summary. Otherwise, the "a" in this model is zero in high probability with the t-test in the summary above.



Q2. ELMR Ch1 Exercise 5.
5. The prostate data — use lpsa as the response.
```{r}
pros = read.table("http://statweb.stanford.edu/~tibs/ElemStatLearn/datasets/prostate.data")
head(pros)
```

```{r}
scatterplotMatrix(pros[,c(1:4,6,9)])

lm <-lm(data = pros, lpsa ~ lcavol)
abline(lm)
ggplot(pros, aes(x = lcavol, y = lpsa)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red")
```
As above I chose lcavol as the x-axis variable to see the response lpsa, based on the linear regression shape in the matrixplot.
As we can see above, from the ggplot function, the linear regreesion red line can be fitted with the x and y variables I wrote.
